A fully labelled proof system for intuitionistic modal logics

### 2020-09-29 FullyLabelled/JLC/conclusion.tex
final pass through paper for now

\review{Also the reason for preferring one semantics for intuitionistic modal logic over another should be discussed in detail, as otherwise the authors’ work would rather look a routine exercise.}
\todo{add a paragraph in conclusion: saying that it is an interesting idea to applly method to CK, but extension is tricky \cite{ADS:lmcs}}

In this paper we embrace the fully labelled approach to intuitionistic modal logic as pioneered by~\cite{maffezioli:etal:synthese13} and generalise it to the class of logics defined by (one-sided intuitionistic) Scott-Lemmon axioms.
%
We establish that it is a valid approach to intuitionistic modal logic by proving soundness and completeness of our system, via a reductive cut-elimination argument.

For a restricted class of logics defined by so-called \emph{path axioms}: $\fm{(\DIA^k\BOX A \IMP\BOX^m A) \AND (\DIA^k A \IMP \BOX^m\DIA A)}$ the standard labelled framework with one relation $R$ was enough for Simpson to get a strong connection between the sequent system, the axiomatisation and the birelational semantics~\cite{simpson:phd}.
%
%\todo{For path axioms we must have that adding the rule for the two sides of the axiom is equivalent to adding the rule that doesn't make use of the pre-order. Is it possible to see that directly on the proof theoretical side?}
%\sonia{probably not because the preorder is introduced/used in other rules for $\BOX$ and $\IMP$}
%
We believe that the framework presented here might be the more appropriate way to treat logics outside of the path axioms definable fragment.

However, we have not showed that our system satisfies Simpson's 6th requirement, that is, "there is an intuitionistically comprehensible explanation of the meaning of the modalities relative to which [our system] is sound and complete".
%
To make sure that his system satisfies this requirement, Simpson chose to depart from the direct correspondence with modal axioms and their corresponding class of Kripke frames, and to study intuitionistic purely as a fragment of intuitionistic first-order logic.
%
We instead took the way of a direct correspondence of our system with the class of frames defined by Scott-Lemmon axioms as uncovered by~\cite{plotkin:stirling:86}, but as this class of logics seems to be rather well-behaved, we believe it shoud be possible to prove the satisfaction of Simpson's 6th requirement too.

>>>>

In this paper we embrace the fully labelled approach to intuitionistic modal logic as pioneered by~\cite{maffezioli:etal:synthese13} and generalise it to the class of logics defined by (one-sided) intuitionistic Scott-Lemmon axioms.
%
We establish that it is a valid approach to intuitionistic modal logic by proving soundness and completeness of our system, via a reductive cut-elimination argument.

For a restricted class of these logics defined by so-called \emph{path axioms}: $\fm{(\DIA^k\BOX A \IMP\BOX^m A) \AND (\DIA^k A \IMP \BOX^m\DIA A)}$ the standard labelled framework with one relation $R$ was enough for Simpson to get a strong connection between the sequent system, the axiomatisation and the birelational semantics~\cite{simpson:phd}.
%
%\todo{For path axioms we must have that adding the rule for the two sides of the axiom is equivalent to adding the rule that doesn't make use of the pre-order. Is it possible to see that directly on the proof theoretical side?}
%\sonia{probably not because the preorder is introduced/used in other rules for $\BOX$ and $\IMP$}
%
We believe that the framework presented here might be the more appropriate way to treat logics outside of the path axioms definable fragment.

However, we have not showed that our system satisfies Simpson's 6th requirement, that is, ``there is an intuitionistically comprehensible explanation of the meaning of the modalities relative to which [the system] is sound and complete".
%
To make sure that his system satisfies this requirement, Simpson chose to depart from the direct correspondence with modal axioms and their corresponding class of Kripke frames, and to study intuitionistic purely as a fragment of intuitionistic first-order logic.
%
We instead took the way of a direct correspondence of our system with the class of frames defined by Scott-Lemmon axioms as uncovered by~\cite{plotkin:stirling:86}, but as this class of logics seems to be rather well-behaved, we believe it should be possible to prove the satisfaction of Simpson's 6th requirement too.

\review{Also the reason for preferring one semantics for intuitionistic modal logic over another should be discussed in detail, as otherwise the authors’ work would rather look a routine exercise.}
\todo{add a paragraph in conclusion: saying that it is an interesting idea to apply method to CK, but extension is tricky \cite{ADS:lmcs}} \sonia{done?}
We have considered in this work the logic known as $\IK$ (and its extensions) with respect to the birelational semantics that is its most well-studied semantics.
%
As we have mentioned, however, different basis can be considered for intuitionistic modal logic, for example the one called \emph{constructive modal logic} $\mathsf{CK}$ and any in between.
%
These can also be studied within the birelational framework with some additional conditions, and we are convinces the fully labelled framework, once extended with these conditions, will be suitable to treat constructive fragments equally.
%
However, treating extensions of those with axioms such as we did here for Scott-Lemmon logics is known to be complex~\cite{ADS:lmcs}.

### 2020-09-29 FullyLabelled/JLC/main.tex
final pass through paper for now

 \begin{abstract}
   \review{Reconsider abstract. It reads very “operational” and does not clearly state the contributions of the paper.}
   \lutz{Done?}
   \sonia{Added two setences at the beginning and modified it a bit.}
   Labelled proof theory has been famously successful for modal logics by mimicking their relational semantics within deductive systems. Simpson in particular designed a framework to study a variety of intuitionistic modal logics integrating a binary relation symbol in the syntax.
   %%
   In this paper, we present a labelled sequent system for
   intuitionistic modal logics such that there is not only one, but
   two relation symbols appearing in sequents: one for the accessibility
   relation associated with the Kripke semantics for normal modal logics and
   one for the preorder relation associated with the Kripke semantics
   for intuitionistic logic.
   %%
   This puts our system in close correspondence with the standard
   birelational Kripke semantics for intuitionistic modal logics. 
   %%
   As a consequence it can be extended with arbitrary intuitionistic Scott-Lemmon axioms. 
   %%
   We show soundness and completeness, together with an internal cut elimination proof, encompassing a wider array of intuitionistic modal logics than any existing labelled systems.
 \end{abstract}

>>>>

 \begin{abstract}
   \review{Reconsider abstract. It reads very “operational” and does not clearly state the contributions of the paper.}
   \lutz{Done?}
   \sonia{Added two sentences at the beginning and modified it a bit.}
   Labelled proof theory has been famously successful for modal logics by mimicking their relational semantics within deductive systems. Simpson in particular designed a framework to study a variety of intuitionistic modal logics integrating a binary relation symbol in the syntax.
   %%
   In this paper, we present a labelled sequent system for
   intuitionistic modal logics such that there is not only one, but
   two relation symbols appearing in sequents: one for the accessibility
   relation associated with the Kripke semantics for normal modal logics and
   one for the preorder relation associated with the Kripke semantics
   for intuitionistic logic.
   %%
   This puts our system in close correspondence with the standard
   birelational Kripke semantics for intuitionistic modal logics. 
   %%
   As a consequence it can be extended with arbitrary intuitionistic Scott-Lemmon axioms. 
   %%
   We show soundness and completeness, together with an internal cut elimination proof, encompassing a wider array of intuitionistic modal logics than any existing labelled systems.
 \end{abstract}
